---
title: "MarkdownNotes"
author: "FedericaFusi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    css: bootstrap.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up the workplace

Upload needed packages
```{r}
library(tidyverse)
library(tidytuesdayR)
```

Open your data
```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 2)
transit_cost <- tuesdata$transit_cost
```

Have a first look at the data
```{r}
transit_cost
```

Check dimensions
```{r}
dim(transit_cost) # 544 rows and 20 columns
```

# Cleaning

## Variables of interest

I want to look at the relationship between **length of the project** and **real cost**. My scope is to *identify those projects whose cost per km is particularly high or low*.

```{r}
class(transit_cost$real_cost) #character
class(transit_cost$length) #numeric
```

The variable **real_cost** needs to be converted into numeric. I quickly look at it to see potential issues that might arise with the conversation. 

```{r}
table(transit_cost$real_cost) #There seems to be a few "weird" entry that are character rather than numeric.
```

I am going to check out these unusual entries by looking at the data
```{r}
View(transit_cost)
```

They seem to be some summary columns left in the dataset by mistake. I will remove them.
```{r}
transit_cost2 =
  transit_cost %>%
  filter(transit_cost$real_cost != "AVG",
         transit_cost$real_cost != "MEDIAN",
         transit_cost$real_cost != "MIN",
         transit_cost$real_cost != "N",
         transit_cost$real_cost != "QUARTILE 1",
         transit_cost$real_cost != "QUARTILE 3",
         transit_cost$real_cost != "STD")

nrow(transit_cost) - nrow(transit_cost2)
```

The pipe has correctly eliminated *7 observations*. The new dataset size is `nrow(transit_cost2)`.

I will check again the column to see if there are any other non-numeric data but everything seems fine.
```{r}
table(transit_cost2$real_cost)
```

I will convert the variable into numeric
```{r}
transit_cost2$real_cost2 = as.numeric(transit_cost2$real_cost)
```

Check missing values in both variables

```{r}
transit_cost2 %>%
  summarise(sum(is.na(real_cost2)),
            sum(is.na(length)))
```

There aren't any, so we are good to start the plot

### A quick note

Note what happens when we convert the original column from the original dataset
```{r}
transit_cost$real_cost2 = as.numeric(transit_cost$real_cost)
```

We get a warning message **Warning message: NAs introduced by coercion**. This a warning message that it is worth checking out. At a minimum you might want to inspect the column manually to see what is going on. 

# My scatterplot

I will start with simple scatterplot to see how observations look like. 

```{r}
transit_cost2 %>%
  ggplot() +
  geom_point(mapping = aes(y = real_cost2, 
                           x = length)) +
  xlab("Length of the line in km") +
  ylab("Total cost in million $") +
  labs(title = "Correlation between lenght and cost by line")
```

There are **seven outliers** whose length of the line is greater than 100 km. These outliers reduce our attention on the vast majority of the project so I decided to remove them. I need to remember to report this information in the graph. 

```{r}
transit_cost2 %>%
  filter(length <=100) %>% # Eliminate the outliers
  ggplot() +
  geom_point(mapping = aes(y = real_cost2, 
                           x = length)) +
  xlab("Length of the line in km") +
  ylab("Total cost in million $") +
  labs(title = "Correlation between lenght and cost by line",
       subtitle = "Only projects below 100 km were included")
```

The graph looks much better now. I want to highlight those projects that are particularly expensive or particularly cheap. I decide to identify them as *being in the top 10 or bottom 10 quantile* considering the cost per km.

```{r}
transit_cost2 %>%
  filter(length <= 100) %>%
  
# I need to create a new variable identify the ratio cost/length
mutate(cost_km = real_cost2 / length,
  
# I now create a new dummy variable indicating whether the ratio is in the top 10 (=2), bottom ten (=0), or in the average (=1)
# I categorize them 0,1,2 so that the label will be ordered as low, average, high cost
         ratio_cost = ifelse(cost_km > quantile(cost_km, 0.90), 2, 
                             ifelse(cost_km < quantile(cost_km, 0.10), 0, 1))) %>%

# Now I start my plot. Same as before but I add color in the aes function to color my dots based on the new cost variable  
  ggplot() +
  geom_point(mapping = aes(y = real_cost2, 
                           x = length, 
                           color = as.character(ratio_cost))) + #Remember to change this as.character or as.factor
  xlab("Length of the line in km") +
  ylab("Total cost in million $") +
  labs(title = "Correlation between lenght and cost by line",
       subtitle = "Only projects below 100 km were included", 
    # I change the legend label
       color = "Project cost per km") +
    # I change the lables of the different category using a scale. Note this is _discrete because I converted the variable into character
  scale_color_discrete(labels = c("Low cost", "Average cost", "High cost"))
 
```

Since I am pretty satisfied by the skeleton of my scatterplot, I can now change the graphic a bit. I want to change the color to highligh low and high cost projects and I want to add more brakes in my y_axis. 

```{r}
transit_cost2 %>%
  filter(length <= 100) %>%
  
  mutate(cost_km = real_cost2 / length,
           ratio_cost = ifelse(cost_km > quantile(cost_km, 0.90), 2, 
                             ifelse(cost_km < quantile(cost_km, 0.10), 0, 1))) %>%
  ggplot() +
  geom_point(mapping = aes(y = real_cost2, 
                           x = length, 
                           fill = as.character(ratio_cost),
                           color = as.character(ratio_cost)), 
             size = 2, 
             shape = 21,
             alpha = 0.5) + 
  xlab("Length of the line in km") +
  ylab("Total cost in million $") +
  labs(title = "Correlation between lenght and cost by line",
       subtitle = "Only projects below 100 km were included", 
       color = "Project cost per km") +
  # Changing color to make
  scale_fill_manual(values = c("#d62828", "#ced4da", "#003049"),
                     labels = c("Low cost", "Average cost", "High cost")) +
  scale_color_manual(values = c("#d62828", "#ced4da", "#003049")) +
  scale_y_continuous(breaks = seq(0, 30000, by = 5000))
```





