---
title: "Assignment 12"
subtitle: "PA 434"
author: "Alexis Kwan"
output: html_notebook
---

```{r include=FALSE}
library(tidyverse)

class(state.name)
class(sentences)
length(state.name)
length(sentences)
```

# Part 1

1. Extract the following states: Arizona, California, Illinois, Oregon
```{r}
str_extract(state.name, "Arizona|California|Illinois|Oregon")
```

2. Extract all states that begin with the letter a. Use both str_subset() and str_extract() to see how
the syntax differs. In both cases, extract the full name of all the states.
```{r}
str_subset(state.name, "^[Aa]")
```
```{r}
str_extract(state.name, "^[Aa].*")
```
`str_subset` seems to return only strings that match the patter while `str_extract` returns the strings that match but in the original positions they appear within the data structure, with everything else being `NA`. 

3. Extract all states whose names finish with a OR e.
```{r}
str_subset(state.name, "[AaEe]$")
```

4. Count how many state names are composed of two words (e.g. North Dakota)
```{r}
str_subset(state.name, ".* .*")
sum(str_count(state.name, ".* .*"))
```

